Raw LendersClub Analysis
========================

Load Libraries
--------------

```{r}
library(maps)
library(Hmisc)
```


Download and Save data
----------------------
Donwload Data, read data in and save it to disk as RDA

```{r loadData,cache=TRUE}
getwd()
download.file("https://spark-public.s3.amazonaws.com/dataanalysis/loansData.csv", destfile="../../data/loansData.csv", method="curl")
dateDownloaded <- date()
dateDownloaded
loansData <- read.csv("../../data/loansData.csv")
save(loansData, dateDownloaded, file="../../data/loansData.rda")
```



### Look at the data set
```{r, dependson="loadData"}
head(loansData)
summary(loansData)
sapply(loansData[1,],class)
```

### Find out about missing values
```{r, dependson="loadData"}
sum(is.na(loansData))
```

### Remove missing value cases
```{r removena, dependson="loadData"}
loansData<- loansData[complete.cases(loansData$Interest.Rate,
                                     loansData$Open.CREDIT.Lines),]
sum(is.na(loansData))
```

### Make Interest Rate , Debt-to-income Ratio, loan Length numeric
```{r, dependson="loadData"}
loansData$Interest.Rate <- as.numeric(gsub("%", "", loansData$Interest.Rate))
loansData$Debt.To.Income.Ratio <- as.numeric(gsub("%", "", loansData$Debt.To.Income.Ratio))
loansData$Loan.Length <-as.numeric(gsub("[^0-9].*", "", loansData$Loan.Length))

str(loansData$Interest.Rate)
```

### Compute representative FICO score
```{r, dependson="removena"}
lowerFICO <- as.numeric(gsub("-.*","", loansData$FICO.Range))
upperFICO <- as.numeric(gsub(".*-", "",loansData$FICO.Range))
loansData$avgFICO <- (lowerFICO+upperFICO)/2
summary(loansData$avgFICO)
```


------

## Exploratory analysis

### Make some univariate plots/summaries

#### Most incomes are below 20K. There are about 2-3 outliers with exceptionally high monthly incomes
```{r, dependson="loadData"}
hist(loansData$Monthly.Income)
```


#### Not a strong relationship between employment length and interest rate
```{r, dependson="loadData"}
boxplot(loansData$Interest.Rate ~ loansData$Employment.Length)
```

#### Loans for Education, renewable energy seems to beget a smaller interest rate
```{r, dependson="loadData"}
plot(loansData$Loan.Purpose , loansData$Interest.Rate, cex.axis=0.65)
```

#### Plot representative FICO Score against interest Rate
```{r, dependson="loadData"}
plot(loansData$avgFICO, loansData$Interest.Rate)
smoothScatter(loansData$avgFICO, loansData$Interest.Rate)
```


### Let's color by salary and by amount requested
#### By Salary
```{r , dependson="loadData"}
salCut <- cut2(loansData$Monthly.Income, g=5)
plot(loansData$avgFICO, loansData$Interest.Rate, col=salCut)
```

#### By Amount
```{r , dependson="loadData"}
amtCut<- cut2(loansData$Amount.Requested, g=5)
plot(loansData$avgFICO, loansData$Interest.Rate, col=amtCut)
```
A lot lot of light blue in the upper region


### Is there a relationship with State
```{r , dependson="loadData"}
boxplot(loansData$Interest.Rate ~ loansData$State)
```


------

## Modeling

### Fit a basic model relating representative FICO to the Interest Rate
```{r lm1, dependson="loadData"}
lm1 <- lm(loansData$Interest.Rate ~ loansData$avgFICO)
summary(lm1)
```

### Plot the regression line on the plot of FICO vs Interest Rate
```{r , dependson="loadData"}
plot(loansData$Interest.Rate ~ loansData$avgFICO)
abline(lm1$coeff[1], lm1$coeff[2])
```

### check residuals vs Observations, and residuls vs fitted
```{r , dependson="lm1"}
par(mfrow=c(1,2))
plot(loansData$avgFICO, lm1$residuals)
plot(lm1$fitted, lm1$residuals)
```

### Lets plot them with color based on amount(see a pattern)
```{r , dependson="lm1"}
par(mfrow=c(1,2))
plot(loansData$avgFICO, lm1$residuals, col=amtCut)
plot(lm1$fitted, lm1$residuals, col=amtCut)
```

### Lets plot them with color based on loan length(see a pattern)
```{r , dependson="lm1"}
par(mfrow=c(1,2))
plot(loansData$avgFICO, lm1$residuals, col=as.factor(loansData$Loan.Length))
plot(lm1$fitted, lm1$residuals, col=as.factor(loansData$Loan.Length))
```

### Let's create new model with these params
```{r , dependson="loadData"}
lm2 <- lm(loansData$Interest.Rate ~ loansData$avgFICO + amtCut + loansData$Loan.Length)
```

### So what does this model look like:
```{r , dependson="loadData"}
summary(lm2)
confint(lm2)

par(mfrow= c(1,3))
smoothScatter(loansData$avgFICO, loansData$Interest.Rate, xlab= "Representative FICO Score", ylab = "Interest Rate (%)")
abline(lm1$coeff[1], lm1$coeff[2])
plot(loansData$avgFICO, lm1$residuals, col= amtCut, xlab= "Representative FICO Score ", ylab = "No adjustment Residuals")
plot(loansData$avgFICO, lm2$residuals, col= amtCut, xlab = "Representative FICO Score", ylab="Full Model Residuals")


```

lm2 <- lm(loansData$Interest.Rate ~ loansData$avgFICO + loansData$Loan.Length + loansData$Amount.Requested + log(loansData$Monthly.Income+1)  + loansData$Amount.Requested)

